Parameterization of the responses of subarctic European vegetation to key environmental variables for ozone risk assessment

### 2021-12-20 review/response-RC1.tex

\item {\color{blue}Line 77. "($\left<[\mathrm{O_3}]\right>$ = 36-54ppb)" Please explain the formalism. What do different brackets mean?} The bracket "$[X]$" notation commonly refers to the concentration of a chemical species X. Parts-per-billion (ppb) is a volume mixing ratio (VMR) and strictly speaking no concentration. Though, they are often used synonymously because measured VMR is reported in concentrations (e.g. $\mathrm{\mu g\,m^{-3}}$) by multiplying with a factor 2. So, perhaps, we should use another terminology instead, e.g. $\chi_{O_3}$ (\href{https://www.e-education.psu.edu/meteo300/node/534}{e-education.psu.edu}). The "$\left<A\right>$" notation is derived from Dirac's "bra-ket" formulism in quantum mechanics, e.g. $\left<\Psi |A|\Psi \right>\rightarrow \left<A\right>$ "this expression gives the expectation value, or mean or average value, of the observable represented by operator A for the physical system in the state $\left|\Psi \right>$ " (\href{https://en.wikipedia.org/wiki/Bra%E2%80%93ket_notation}{wikipedia}).

\item {\color{blue}Line 81. "A substantial body of evidence exists that suggests flux-based metrics, that relate stomatal ozone uptake to vegetation damage, are biologically more relevant for risk assessments than exposure-based metrics." 
Well, please cite at least some works of this "substantial body"} Thank you, we will include relevant citations.

\item {\color{blue}Figure 1. This figure was never referred in the text.}
This Figure has been referred to in Sect.~2 L177/179.

\item {\color{blue}Figure 2. It does not seem to me that the O3 concentrations of 2019 are different from those of 2018. The spring peak could even be identical (although unknown, because in 2019 O3 measurements started about 20 days after the spring peak)}
In Sect.~2, we assess in detail whether and how those two years differ. We will make this analysis more comprehensible and comprehensive in the next revision.\\
Regarding the spring peak, until the beginning of May most of the vegetation in the subarctic is still either covered by snow (see Table~3 for the dates) or bud burst has not yet occurred. Hence, it was unlikely for deciduous vegetation and scrubs or other short vegetation to experience the ozone spring peak in 2018/19. We will come back to coniferous trees later.

@@

\item {\color{blue}Line 146. Please make clear the acronym PFT on first use}
Plant functional type (PFT) was introduced in the abstract. Following the common code of conduct, we ought to introduce it also in the main matter. Thank you for pointing this out.

\item {\color{blue}Line 165. "luftkvalitet.no" What is it? And EBAS? Please make them clear.}
"luftkvalitet.no" is the webinterface to the database of air quality measurements in Norway operated by NILU. EBAS is a database with webinterface operated by NILU as well. It comprises global air quality measurements, e.g. trace gases, pollutants, particular matter from different networks. We will clarify this.

@@

\item {\color{blue}Line 182. "This indicates that the vegetation was more affected by ozone in 2018 than in 2019." Being affected by visible symptoms does not necessarily mean having suffered biomass or productivity reduction.} It is correct that biomass reduction and visible damage due to ozone are not the same. We discuss this briefly in Sec.~5 but will elaborate on it to make the difference clearer.

\item {\color{blue}Line 190. "high ozone concentrations ([O3] $>$ 40ppb" It is strange to read that O3 concentrations above 40 ppb are "high" concentrations.}
Indeed 40\,ppb is not particularly "high" compared to peak ozone pollution downwind of densely populated areas, e.g. in Eastern Asia. We will remove the term "high". However, ozone concentrations of $40\,\mathrm{ppb}$ have been chosen as critical level for the metric accumulated over a threshold (AOT) to gage ozone exposure toxic to ecosystems. This metric was first introduced at the UN/ECE-Workshop in Egham in 1992. For the protection of crops and semi-natural vegetation (limit to 5\% yield reduction) and forest different targets exist (see e.g. \href{https://www.eea.europa.eu/publications/TOP08-98/page008.html}{EU directive}).

@@

\item {\color{blue}Line 194. "A method for gapfilling data has been presented in Falk et al. (2021)."
Ok, but was it then applied to this work? Please write it.}
We will rephrase: "A method for gapfilling data has been presented in Falk et al. (2021) and applied in this work."

\item {\color{blue}Line 205. "We evaluate the statistical significance of divergences from the norm in these variables (referred to as anomalies) in 2018/19"
I suspect a misuse of the locution "statistical significance". How was this significance assessed? Which statistical test was applied? What is the level of significance?}
We will elaborate on the hypothesis test in the next revision and may show the residuals as violin plots to increase the transparency of our method and make the data more assessable to the reader.\\
In the recent version of the manuscript, we computed the excess number of days with residuals larger than $1\sigma$ first. This means, we calculate the daily mean and std for 21 years $\times$ 24 hours (264 data points).
From this, we find the residuals in 2018/19 and divide by std. In the next step, we explicitly assume that these are normal distributed - we will elaborate on this assumption in the next revision and check the underlying distributions (but see also Fig.~4). Due to natural variability (diurnal, annual) in the data, we chose a $1\sigma$ level for the significance test. Because we are not looking at repeated experiments with well constraint conditions, a much higher sigma level is not necessarily meaning full. But the diurnal cycle could be perhaps removed to smooth the data. Second, by grouping the data into months, we find the number of days with observations in each month. We count the number of days with residuals exceeding $1\sigma$ and compute the ratio with the number of days of observation. If this data was normal distributed, we would expect ~18\% of data in this interval. If the percentage of days exceeding $1\sigma$ is larger, we state that we find a significant excess of days with high ozone.

\item {\color{blue}Lines 206-208. I do not understand. Please, explicit the methodology.}
We will rephrase and include a brief description of the method. We use a Reynolds decomposition on the 1986-1996 data from Svanvik and data from all long-term monitoring stations in Fennoscandia (1992-2012, Jergul-Karasjok, Pallas, Esrange) to establish multi-annual means. We then remove the annual cycle from the July 2018 observations at Pallas (reference station) to find the anomalies. We have shown that these correlate best with observations at Svanvik. We find the time lag between the observations and shift the time series of anomalies accordingly. We then add the climatology of Svanvik to retrieve a reconstruction of the missing data.\\
As 2019 does not deviated much from the norm by means of temperature, precipitation, and irradiance, we assume it for a normal year. From this, we find a probable offset of the historical ozone climatology compared to present day and correct for this. We compared this reconstruction with high resolution air quality model reanalysis and found a similar accuracy compared to observation. For details, please refer the cited work Falk et al (2021).

@@

\item {\color{blue}Line 213. "Averaged monthly accumulated precipitation (blue bars) is shown with standard deviation" It is not consistent to show SE once and STDEV the other time. The use of SE is more appropriate when estimating averages.}
We will change the figure accordingly.

\item {\color{blue}Line 228. "Darker colors indicate higher probability to observe these values."\\Line 229. "On top of the density distributions, a 10 days average of daily mean (h[O3]i10d) is displayed together with 1sigma uncertainties and SE, respectively" What does it mean? It is not clear to me. Why show a probability density if you are plotting a multiannual average? Or does the line represent the median instead?} In Fig.~4, we show the distribution of ozone data in 2-dimensions. The dotes represent the daily averages (but only every 10th day is shown). We will rephrase the caption to correct for this. We show the daily average to infer that not all daily mean ozone are strictly normal distributed (tails in one or the other direction) but it is also visually deductible that most data fall in the $1\sigma$ band justifying the chosen confidence intervals for a normal distribution. We shall, however, elaborate on this by statistical means.

\item {\color{blue}Line 232. "The decline in h[O3]i coincides with the average beginning of CO2 uptake by coniferous trees (Kolari et al., 2007; Wallin et al., 2013)" I didn't know that evergreens only uptake CO2 starting in May. I was convinced they always did. Is it true? Doesn't that contradict what you wrote in line 309 ("We base our temperature acclimation of coniferous trees on experimental results on Norway spruce which were found to be active already at rather low air temperatures and can reach 60\% photosynthetic activity as early as doy 100 (Kolariet al., 2007; Wallin et al., 2013).")? Here you state that photosynthesis is already active at DOY 100 and is at 60\% of its maximum!}
It is true that coniferous trees can, in principle, take up $\mathrm{CO_2}$ also in winter and the fact that they can suffer from something called frost drought shows that. With respect to the spring peak that was mentioned earlier, this means that coniferous trees may experience these elevated ozone concentrations regularly. Concerning the contradiction in the statements -- if the conditions are right, coniferous trees can quickly reach 60\% of their max photosynthesis but stop photosynthesis the days after. Conditions to keep up this levels of photosynthesis, however, would only occur from doy 100 onwards. To make sure that we do not falsely accumulate too much ozone from the spring peak periode (when coniferous trees show on-off photosynthesis), we consulted satellite observations to extrapolate the dates when vegetations' photosynthesis (including coniferous trees) at Svanhovd rose over the detectable threshold.

### 2021-12-20 review/response-RC3.tex

% Authors' response
\documentclass{scrartcl}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{multirow}
\thispagestyle{empty}

\begin{document}
\section*{Authors' response}
To bg-2021-260-RC3 (30 Nov 2021):
We thank the anonymous referee \#3 for their comments.
We will address all \textbf{general comments} raised regarding the structure and readability of the manuscript in our next revision or resubmission. We consider to separate the manuscript in Introduction--Methods--Results--Conclusions and will improve on the "methods" Section to make our results more comprehensible. We will refine the language to reduce ambiguity and supplement our statements with appropriate citations where indicated by the referee.\\
As the manuscript will, hence, undergo major structural changes, we will not address all \textbf{specific / technical comments} in detail at this point. We will take all specific and technical comments which are still relevant into consideration.\\
In the following, we shall give a brief respond to all relevant issues.

\subsection*{Major comments} 
\begin{itemize}

    \item {\color{blue}Concerning the length of the paper, I really cannot see the benefit of all the work on long-
    term ozone and meteorology given in Sections 2-3. Isn't the main point of this paper to
    illustrate differences in vegetation characteristics between plants at nearly 70 degrees N
    and the Mapping Manual (MM)? That doesn't need a long analysis of climate differences or
    ozone trends.}

    \item {\color{blue}Concerning the "default" growing seasons (from DOY 100 to 307) from the MM against
    which the "bespoke" are compared, these are just wrong for this latitude. The MM
    suggests a latitude function that gives DOY 129 to 268.}

    \item {\color{blue}Some of the basic equations look wrong to me, see my comments on Appendix B below.}

    \item {\color{blue}On p21 the author's first "key findings" state that their "bespoke parameterizations for
    subarctic species are better suited..." than the MM ones. This is a strong statement given
    that the authors admit on L293 that their bespoke suggestions are "hypothetical and have
    yet to be verified by observations".}

### 2021-12-20 review/response-RC3.tex

    \item {\color{blue}On p21 the author's first "key findings" state that their "bespoke parameterizations for
    subarctic species are better suited..." than the MM ones. This is a strong statement given
    that the authors admit on L293 that their bespoke suggestions are "hypothetical and have
    yet to be verified by observations".}

\end{itemize}

\subsection*{Other comments} 
\begin{itemize}

\item {\color{blue}}

\item {\color{blue}L14. one can ask if an underestimate of 6\% is a significant problem. I would guess that the overall uncertainty of POD calculations is far greater than this!}

\item {\color{blue}L22. Better to say "estimated" rather than reported. These numbers do not represent actual measured yield losses. Also, as this paper is mainly discussing DO3SE-type approaches to POD, it would have been relevant to cite the DO3SE-based estimates of the Mills et al 2018 paper (below) which also made yield estimates for four crops.}

\item {\color{blue}L29. A more recent citation would be Tarasick et al 2019.}

\item {\color{blue}L36. Are these numbers for tropospheric ozone relevant to this paper? Vegetation is affected by near-surface ozone, and levels of this are much lower than 50-65 ppb over the Nordic countries.}

\item {\color{blue}L39-40. Why "despite"? 22 days is a very long time! This time-scale is not so relevant though; it is the lifetime of ozone in the boundary layer that matters, and that is about 1 week (Hov et al., 1978, or as can be estimated by typical deposition velocities and boundary layer depths). Solberg et al. 2008 also show how residence times of less than 7 days are usual for ozone episodes.}

\item {\color{blue}L48. Spring peaks are seen across large swathes of the northern hemisphere, for example at Mace Head, caused by many factors (Monks et al., 2009). In fact, Monks et al cite Winkler (1988) that north of 60N, the maximum in ozone reduces in amplitude and shifts to later months.}

\item {\color{blue}L85. I don't think Simpson et al 2007 or Mills et al 2011, 2017 focused on the Mediterranean? These were pan-European studies, that also stressed that POD as a metric showed that risks in e.g. northern Europe were likely higher than the earlier AOT40 metrics showed. The ICP-vegetation type metrics have in fact been based upon data from countries across Europe, and with more data from central and northern than from southern European countries.}

\item {\color{blue}L89. Continuing the above, the studies from Karlsson et al have had a strong focus on Scandinavian vegetation.}

\item {\color{blue}L96. Define PLA}

\item {\color{blue}L97. I don't think you need this about typically exceeded under daylight hours. POD calculations follow DO3SE methods, and the flux is zero at night by definition.}

\item {\color{blue}L103. The Maas et al book isn't so easily available (and no web-access address is given anyway). It can be good to cite some of the other key papers behind ozone CLs, e.g. Fuhrer et al. (1997), Mills et al 2011 (below).}

\item {\color{blue}L105. "The CL is calculated by"? Do you mean exceedance of CL?}

\item {\color{blue}L110. "For our study...". This confused me. Are the authors reporting results from the present manuscript, or from some unpublished and uncited study? The sentences starting here seem to be out of place.}

\item {\color{blue}L127. Do you mean non-methane VOC (NMVOC)?.}

\item {\color{blue}L128. Why brackets in "[O3]"?}

\item {\color{blue}L131-132. I didn't quite understand "shall serve as a reference". Do you mean as an example or possible future?}

\item {\color{blue}L141. Why "sensu" here. I had to look up the word, and read that it is "used especially in technical taxonomic references". Use simpler English.}

\item {\color{blue}L141. In any case what does "(sensu World Meteorological Organisation)" mean in this sentence?}

\item {\color{blue}L165-167. Mangled sentences? Why "Last accessed"? There is no url here. Placement of (a) and (b) is strange.}

\item {\color{blue}L170-171. Why deal with data that may introduce a false trend at all? Why wouldn't the problems influence seasonality? I would guess that weather and climate conditions may well have influenced the frequency of quality-assurance checks.}

\item {\color{blue}L172. Here it says that agromet variables are available from 1992, so, again, why use problematic O3 data from the 1980s?}

\item {\color{blue}L214, Fig. 3. The stars are very small on the figure. Make these more obvious. Or just skip them - doesn't the temperature curve give enough information?}

\item {\color{blue}L215-224. Do we need to be told all of these details with plus/minus limits? These are just climatalogical values and can be seen well enough in Fig. 3. The numbers are not used for anything.}

\item {\color{blue}L255-256. Why "deduce"? Now the numbers given, 800 W/m2 and 200 W/m2, are very rounded - are these measurements or assumptions or limits?
L276. I would say hypothetical or local rather than bespoke.}

\item {\color{blue}L280. Again, why "deduce"?}

\item {\color{blue}L281-282. The better ref here for the generic PFTs is the 2017 Mapping Manual. Or what did you use from Simpson et al 2007 or Mills et al. 2011 that isn't in Mills et al 2017?}

\item {\color{blue}L286. First, what is Gstoleaf compared to gsto? On L553 it is stated that Gsto(leaf) is at leaf-level, which suggests that same as gsto in eqn (B1). Then, you claim that Fig. B1 shows low Gsto(leaf), but it shows enormous values! According to Fig. B1, Gstoleaf is ca. 100 mmole/m2/s, this is 1.0e5 nmole/m2/s! This makes no sense.}

\item {\color{blue}L292. The authors say that their fT system has not been verified, but haven't ecosystem models (e.g. JULES, CLM, LPJ-GUESS) parameterized such vegetation already? I would have thought that there was something to learn from such models.}

\item {\color{blue}L344. As noted above, why compare your Astart (129, 130) with the MM value of 100 which is valid for 50 degrees N? The MM suggests a latitude function, which gives a start date of 129 for your location - i.e. almost perfect, and your comment was misleading. Why would you consider an agricultural criteria for forests anyway? Why not apply the MODIS response for deciduous trees too?}

\item {\color{blue}Table 3: Again, the "Defaults" given here are incorrect for this latitude. The correct ref should anyway be the MM.}

\item {\color{blue}L345. "Due to the lack of quantitative field observation...". I didn't follow the logic here. Table 4. Why average the MM but not the bespoke? Give the MM for both years - this is useful information.}

\item {\color{blue}L488. Give references for your statements that global land-surface models have problems in the Arctic regions.}

\item {\color{blue}L624-625. Mangled ref. And why source googlebooks, when this report is easily available from www.emep.int?}

\item {\color{blue}References: check upper-lower case, in e.g. AMBIO, grennfelt etc.}

\item {\color{blue}Appendix B: DO3SE model
The equations used here, and especially around L524-532, have several problems:
\begin{itemize}
\item they do not account for differences between canopy and leaf-scale conductances
\item they make no mention of how the ozone is calculated at the height of the vegetation from the 2m measured values.
\item Equation B6 looks wrong. What is the u(z1) term doing here?
\item What is "z1" in eqn. B7?
\item L531 the factor 1.3 accounts for the "difference in" diffusivity.
\item Figure B2 - does MODIS only respond to coniferous trees in this 1x1 km2 grid? No other vegetation?
\end{itemize}}

### 2022-01-05 review/response-RC3.tex

% Authors' response
\documentclass{scrartcl}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{multirow}
\thispagestyle{empty}

\begin{document}
\section*{Authors' response}
To bg-2021-260-RC3 (30 Nov 2021):
We thank the anonymous referee \#3 for their comments.
We will address all \textbf{general comments} raised regarding the structure and readability of the manuscript in our next revision or resubmission. We consider to separate the manuscript in Introduction--Methods--Results--Conclusions and will improve on the "methods" Section to make our results more comprehensible. We will refine the language to reduce ambiguity and supplement our statements with appropriate citations where indicated by the referee.\\
As the manuscript will, hence, undergo major structural changes, we will not address all \textbf{specific / technical comments} in detail at this point. We will take all specific and technical comments which are still relevant into consideration.\\
In the following, we shall give a brief respond to all relevant issues.

\subsection*{Major comments} 
\begin{itemize}

    \item {\color{blue}Concerning the length of the paper, I really cannot see the benefit of all the work on long-term ozone and meteorology given in Sections 2-3. Isn't the main point of this paper to illustrate differences in vegetation characteristics between plants at nearly 70 degrees N and the Mapping Manual (MM)? That doesn't need a long analysis of climate differences or ozone trends.} A detailed analysis of long-term ozone and meteorological conditions is important for to modeling of ozone risk in the subarctic. Our sample period only extends over two growing seasons (2018/2019). To assess wether or not the visible damage could correspond to an unusual high ozone dose in 2018, we need to set this particular year into a climatological context. Upon restructuring the manuscript, we may remove or condense this part.

    \item {\color{blue}Concerning the "default" growing seasons (from DOY 100 to 307) from the MM against which the "bespoke" are compared, these are just wrong for this latitude. The MM suggests a latitude function that gives DOY 129 to 268.} The referee is right regarding the latitude model at $70\,^\circ$N and we could have used these instead from the start. However, as we state in L344, these start and end DOYs for grassland refer to central Europe.

    \item {\color{blue}Some of the basic equations look wrong to me, see my comments on Appendix B below.} We will double check the equations make the description of the $\mathrm{DO_3SE}$ model more comprehensible and comprehensive. However, in the interest of the main focus of this manuscript it may not be possible to give a full recap of all details concerning this well established model.

    \item {\color{blue}On p21 the author's first "key findings" state that their "bespoke parameterizations for subarctic species are better suited..." than the MM ones. This is a strong statement given that the authors admit on L293 that their bespoke suggestions are "hypothetical and have yet to be verified by observations".} Thank you for pointing this out. We will rephrase our key findings accordingly.

### 2022-01-05 review/response-RC1.tex

\item {\color{blue}Line 336. Using net photosynthesis to calculate leaf emergence is not completely justified. Leaves are likely to be present and active well before gross photosynthesis equals heterotrophic respiration (eg. soil respiration). Gross photosynthesis should be used to calculate Astart and Aend instead.}
Thank you for pointing this out. It is correct that net photosynthesis does not capture the real photosynthesis. However, gross photosynthesis is not among the available products from MODIS satellites. Instead, we use gross primary production which is including maintenance respiration. According to the product website (\href{http://www.ntsg.umt.edu/project/modis/mod17.php}{MOD17}, last accessed Dec. 2021) it is computed from observed absorbed photosynthetically active radiation (APAR), a species specific conversion efficiency parameter $\epsilon(\mathrm{temperature, water stress})$, temperature, and water stress. In Fig.~\ref{fig:modis_gpp} the estimated start and end of growing season from MODIS (Aqua/Terra) GPP over a $1\times 1$\,km patch centered at Svanhovd is shown. They differ only slightly from the estimates using net photosynthesis. The start of the growing season is shifted by one day (later) and the end by one/two days (later). Considering the temporal resolution of 8\,days of the satellite product this difference is negligible.

\begin{figure}[!th]
    \centering
    \includegraphics[width=0.75\textwidth]{./modis_Gpp.png}
    \caption{Estimated start and end of growing season from MODIS (Aqua/Terra) GPP over a $1\times 1$\,km patch centered at Svanhovd. The average start in 2018 amounts to doy 123 and to 107 in 2019. The end of the growing season in 2018 is doy 262 and in 2019 276.}
    \label{fig:modis_gpp}
\end{figure}

\item {\color{blue}Line 354. "A sample of downy birch leaves collected at Svanhovd had an average length of (3.0±0.5)cm" Were top-canopy leaves sampled? How many leaves were collected to get +- 0.5 cm standard error?}
The birch (B. pubescence) leaves were collected from the outer canopy (good light exposure) of tall trees at two sites in Finnmark. The leaves were collected late in July so they were fully expanded. They were collected by hand, so at about $2\,\mathrm{m}$ height. As the sun is at a low elevation angle for most of the time in the growing season in this area, light exposure at the top of the tree is not expected to be so different from lower leaves as long as they grow in a part of the tree that is not shadowed by other trees. At the first site (Karasjok), three leaves were collected from each of five adult trees. At the second site (Svanvik), five leaves were collected from one tree. The leaves were pressed and dried before scanning for the area and shape determination through image analysis. ImageJ was used for thresholding the images to give silhouettes of the leaves. After scaling the images, the program was used for finding minimum Feret diameter. This measure finds the largest width of a leaf at a $90^\circ$ angle to the length of the leaf (the Feret diameter). With this method we found that the 20 leaves had a mean width of $3.014\,\mathrm{cm}$ and a standard deviation of $0.4996\,\mathrm{cm}$, given as $(3.0\pm0.5)\,\mathrm{cm}$ in the manuscript.

### 2022-01-05 review/response-RC3.tex

% Authors' response
\documentclass{scrartcl}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{multirow}
\thispagestyle{empty}

\begin{document}
\section*{Authors' response}
To bg-2021-260-RC3 (30 Nov 2021):
We thank the anonymous referee \#3 for their comments.
We will address all \textbf{general comments} raised regarding the structure and readability of the manuscript in our next revision or resubmission. We consider to separate the manuscript in Introduction--Methods--Results--Conclusions and will improve on the "methods" Section to make our results more comprehensible. We will refine the language to reduce ambiguity and supplement our statements with appropriate citations where indicated by the referee.\\
As the manuscript will, hence, undergo major structural changes, we will not address all \textbf{specific / technical comments} in detail at this point. We will take all specific and technical comments which are still relevant into consideration.\\
In the following, we shall give a brief respond to all relevant issues.

\subsection*{Major comments} 
\begin{itemize}

    \item {\color{blue}Concerning the length of the paper, I really cannot see the benefit of all the work on long-term ozone and meteorology given in Sections 2-3. Isn't the main point of this paper to illustrate differences in vegetation characteristics between plants at nearly 70 degrees N and the Mapping Manual (MM)? That doesn't need a long analysis of climate differences or ozone trends.} A detailed analysis of long-term ozone and meteorological conditions is important for to modeling and understanding of past, present, and future ozone risk in the subarctic. Our sample period only extends over two growing seasons (2018/2019). To assess wether or not the visible damage that we observed in 2018 could correspond to an unusual high ozone dose, we had to to set this particular year into a climatological context. Upon restructuring the manuscript, we may condense this part.

@@

    \item {\color{blue}On p21 the author's first "key findings" state that their "bespoke parameterizations for subarctic species are better suited..." than the MM ones. This is a strong statement given that the authors admit on L293 that their bespoke suggestions are "hypothetical and have yet to be verified by observations".} Thank you for pointing this out. We will rephrase our key findings accordingly.

\end{itemize}

\subsection*{Other comments} 
\begin{itemize}

\item {\color{blue}L14. one can ask if an underestimate of 6\% is a significant problem. I would guess that the overall uncertainty of POD calculations is far greater than this!} To assess the underlying uncertainty caused by choice of the parameterizations that are used to compute POD, we have varied the temperature and light response paramterizations and looked at the influence of SWP in our target region. Within this parameter space the uncertainty due to temperature and light response is $\approx3\,\%$ for deciduous trees. (For which we found the $6\,\%$ underestimate.) Hence, we conclude that the estimated underestimate is significant (albeit not in a strict statistical sense).

\item {\color{blue}L22. Better to say "estimated" rather than reported. These numbers do not represent actual measured yield losses. Also, as this paper is mainly discussing DO3SE-type approaches to POD, it would have been relevant to cite the DO3SE-based estimates of the Mills et al 2018 paper (below) which also made yield estimates for four crops.} Thank you for pointing this out. We will rephrase and include the relevant citations.

\item {\color{blue}L29. A more recent citation would be Tarasick et al 2019.}
Tanks you. We will include the citation.

\item {\color{blue}L36. Are these numbers for tropospheric ozone relevant to this paper? Vegetation is affected by near-surface ozone, and levels of this are much lower than 50-65 ppb over the Nordic countries.}
These numbers are indeed not relevant for the monitoring station at Svanhovd which is situated at low altitude. We will remove the reference.

\item {\color{blue}L39-40. Why "despite"? 22 days is a very long time! This time-scale is not so relevant though; it is the lifetime of ozone in the boundary layer that matters, and that is about 1 week (Hov et al., 1978, or as can be estimated by typical deposition velocities and boundary layer depths). Solberg et al. 2008 also show how residence times of less than 7 days are usual for ozone episodes.} Thank you for pointing this out. We will rephrase the sentence to include the more relevant time scales. Wehen citing 22 days, we had the long-range transport (trans-continental) in mind.

\item {\color{blue}L48. Spring peaks are seen across large swathes of the northern hemisphere, for example at Mace Head, caused by many factors (Monks et al., 2009). In fact, Monks et al cite Winkler (1988) that north of 60N, the maximum in ozone reduces in amplitude and shifts to later months.}
Thank you for pointing this out. We will rephrase the sentence accordingly.

\item {\color{blue}L85. I don't think Simpson et al 2007 or Mills et al 2011, 2017 focused on the Mediterranean? These were pan-European studies, that also stressed that POD as a metric showed that risks in e.g. northern Europe were likely higher than the earlier AOT40 metrics showed. The ICP-vegetation type metrics have in fact been based upon data from countries across Europe, and with more data from central and northern than from southern European countries.}
We will carefully check the cited works and rephrase the paragraph accordingly.

\item {\color{blue}L89. Continuing the above, the studies from Karlsson et al have had a strong focus on Scandinavian vegetation.}
See above.

\item {\color{blue}L96. Define PLA}
We will include the definition "per leave area" in the revision.

\item {\color{blue}L97. I don't think you need this about typically exceeded under daylight hours. POD calculations follow DO3SE methods, and the flux is zero at night by definition.} We will remove it accordingly.

\item {\color{blue}L103. The Maas et al book isn't so easily available (and no web-access address is given anyway). It can be good to cite some of the other key papers behind ozone CLs, e.g. Fuhrer et al. (1997), Mills et al 2011 (below).} We will include the given references in addition.

\item {\color{blue}L105. "The CL is calculated by"? Do you mean exceedance of CL?} Of course, we mean the exeedance.

\item {\color{blue}L110. "For our study...". This confused me. Are the authors reporting results from the present manuscript, or from some unpublished and uncited study? The sentences starting here seem to be out of place.}
The ozone monitoring at Svanhovd was originally meant as a supplement for another study in our project. Hence, the confusing formulation. We will rephrase the sentence.

\item {\color{blue}L127. Do you mean non-methane VOC (NMVOC)?.} Yes, we mean non-methane VOCs.

\item {\color{blue}L128. Why brackets in "[O3]"?}
The bracket "$[X]$" notation commonly refers to the concentration of a chemical species X. Parts-per-billion (ppb) is a volume mixing ratio (VMR) and strictly speaking no concentration. Though, they are often used synonymously because measured VMR is reported in concentrations (e.g. $\mathrm{\mu g\,m^{-3}}$) by multiplying with a factor 2. So, perhaps, we should use another terminology instead, e.g. $\chi_{O_3}$ (\href{https://www.e-education.psu.edu/meteo300/node/534}{e-education.psu.edu}).

\item {\color{blue}L131-132. I didn't quite understand "shall serve as a reference". Do you mean as an example or possible future?}
Exactly. "Example" may be the right term in this context.

\item {\color{blue}L141. Why "sensu" here. I had to look up the word, and read that it is "used especially in technical taxonomic references". Use simpler English.}
We will rephrase.

\item {\color{blue}L141. In any case what does "(sensu World Meteorological Organisation)" mean in this sentence?}
In general it means "in the sense of". But in this context, it could also be read as "refer to" for further reading. "Essential climate Variables" is a term defined and commonly used by the WMO that might not be well known in the plant physiology community.

\item {\color{blue}L165-167. Mangled sentences? Why "Last accessed"? There is no url here. Placement of (a) and (b) is strange.}
This was supposed to be in the bibliography. We will check the bibtex citation styles to fix the (a) and (b) placements.

\item {\color{blue}L170-171. Why deal with data that may introduce a false trend at all? Why wouldn't the problems influence seasonality? I would guess that weather and climate conditions may well have influenced the frequency of quality-assurance checks.\\
L172. Here it says that agromet variables are available from 1992, so, again, why use problematic O3 data from the 1980s?}
These two questions are related. Consulting the ozone monitoring report by Solberg et al. (2003), we find an inaccuracy in the text which should read "[...] from before 1997 [...]". The systematic uncertainty in the pre 1997 ozone date is of the order of $10\,\%$ in AOT. Which means that the uncertainty in each individual measurement is lower but in the order of the trend in background ozone. This concerns the whole historical monitoring period at Svanhovd. Combining all data results in a fair confidence in the observed seasonality rather than each individual measurement. From this we find, that 2019 was a rather usual year in terms of ozone and weather. Stressing that the $6\,\%$ underestimation of biomass loss for deciduous trees is rather the norm than an exception. Without relaying on an ozone climatology, we could not make this assessment.

\item {\color{blue}L214, Fig. 3. The stars are very small on the figure. Make these more obvious. Or just skip them - doesn't the temperature curve give enough information?}
We will revise the figure accordingly.

\item {\color{blue}L215-224. Do we need to be told all of these details with plus/minus limits? These are just climatalogical values and can be seen well enough in Fig. 3. The numbers are not used for anything.}
We consider removing these in the next revision.

\item {\color{blue}L255-256. Why "deduce"? Now the numbers given, 800 W/m2 and 200 W/m2, are very rounded - are these measurements or assumptions or limits?}
We assume the referee refers to L225-226? We agree that "deduce" is not the right term in this context and will rephrase. These are limits based on the observations.

\item {\color{blue}L276. I would say hypothetical or local rather than bespoke.}
Thank you for pointing this out. We will change the term accordingly.

\item {\color{blue}L280. Again, why "deduce"?}
See above.

### 2022-01-05 review/response-RC1.tex

\item {\color{blue}Figure A1. How was the length of the growing seasons in the various years identified? By satellite? Other method? What does the gray band represent?} See response to "Line~432". The gray band represents the standard deviation. We will either remove or elaborate on this analysis.

\item {\color{blue}Line 511. "with fmin, Dmin, Dmax describing the relative stomatal conductance to changes in vapor pressure deficit." It is not clear. Please, clarify what D and fmin are, and their units.}

\item {\color{blue}
\begin{itemize}
    \item Line 517. "The DO3SE model as described in Büker et al. (2012) is used to simulate SWP0 across a PFT specific root depth according to the Penman--Monteith energy balance method that drives water cycling through the soil--plant--atmosphere system" I cannot understand how the P-M energy balance is used in DO3SE to derive the SWP. Please explain in detail.
    {\color{black} This is described in Bücker et al (2012) in detail. We will include an improved, short description of how the $\mathrm{DO_3SE}$ model incorporates the PM model to estimate evapotranspiration and hence water loss from the soil. However, for full details of this method it is best for readers to refer to Bücker et al. (2012).}

    \item Line 525. "the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf" At what height was the O3 concentration measured? If it was not measured at the top of the canopy (10m for trees or 10 cm for grassland), how was the O3 concentration at the top canopy calculated? Please explain in detail.
    {\color{black}The $\mathrm{DO_3SE}$ model was used to estimate the difference in $[\mathrm{O_3}]$ between a reference height above the canopy and the canopy height. This employs the deposition component of the $\mathrm{DO_3SE}$ model that estimates the transfer of mass (i.e. ozone) as a function of windspeed, convection, surface roughness, and vegetation $[\mathrm{O_3}]$ sink. We will add additional detail in the paper to make this clear.}
